On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-06-27 Manuscript/manuscript.tex
added graph
Started responding to prompts and breaking Rank comparison into 2 smaller subsections

\subsection{World Championships}\label{sec:dataworld}

@@

\subsection{Beyond World Championships}\label{sec:databeyond}
The motivation for looking beyond World Championships data came from an 
exploratory analysis examining how United States athletes reacted at the 
2022 United States Track and Field Championships.

From June 23-26 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World Championships 
being held at the same venue in August. Thus, we can establish a baseline for the four United States athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championships races compared to the United States Track and Field 
Championships races. Devon Allen raced three times at the United States Track and
Field Championships and three times at the World Track and Field Championships. 
His reaction times were substantially faster in the World Track and
Field Championships compared to the United States Track and Field Championships, 
as were his
teammates. 
This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction 
times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts.

@@

  \end{center}
\end{table}

\eds{Can you show some descriptive statistics for this data?  Total number of 
clusters? Average number of international vs national competitions across 
athletes/clusters?}

\eds{The following paragraph still needs to be smoothed (the paragraph that was 
here and the paragraph that was in the intro still need to be integrated into 
one (or two) cohesive paragraph(s) that accurately describe the two types of 
data that will be used for two rank-based analyses.}
The dataset of the four United States athletes 
is too small to perform any meaningful statistical analysis, but
the data was expanded to include reaction times for athletes who competed at 
national competitions (many of which were held from May-July of 2022). Once the 
data was collected, each athlete was designated as a cluster and the stage of
competition (national or international) was specified for each competition 
within the cluster (athlete). First we can compare the reaction times of 
athletes who
competed at the World Track and Field Championships in both 2019 and 2022, and 
then we can compare how athletes performed at the 2022 championships relative 
to their national/qualifiers meet. In order to reduce any chances
of athletes improving reaction time, we can also look at data from qualifying 
meets. Almost every athlete had to qualify by proving they are elite at the 
national level, usually by coming in at least fourth. Thus, data was compiled on 
athletes who competed at the 2022 World Track and Field Championships on what 
these athletes also ran at their country's national championship. If there was 
any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August. The equipment used to measure 
reaction time cannot be guaranteed to be the same from national to international
competition, but the time span between meets has been reduced considerably.

\begin{figure}[tbp]
  \centering
  \includegraphics{2019vs2022BarGraph}
  \caption{The reaction times from 2019 and 2022 World Track and Field
  Championship. ``H'', ``S'', and``F'' refer to the heats, semifinals, and finals
  respectively.}
  \label{fig:2019vs2022Graph}
\end{figure}
\of{I included this here for now, but feel free to change the location.  I will
add references to it and explanation of it along with some descriptive statistics about
the size and mean/median,std dev, size of the data.}
\jy{Show dot plots which would give the distribution; without uncertainty
  measure, the mean does not tell much.}

>>>>

\subsection{2022 World Championships}\label{sec:dataworld}

@@

\subsection{Beyond 2022 World Championships}\label{sec:databeyond}
The motivation for looking beyond World Championships data came from an 
exploratory analysis examining how United States athletes reacted at the 
2022 United States Track and Field Championships.

\subsection{2022 National Competitions}\label{sec:datanational}
From June 23-26 2022, the United States held its Track and Field Championships 
at Hayward Field in Eugene, Oregon to decide who to send to the World Championships 
being held at the same venue in August. Thus, we can establish a baseline for the four United States athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
Track and Field Championships races compared to the United States Track and Field 
Championships races. Devon Allen raced three times at the United States Track and
Field Championships and three times at the World Track and Field Championships. 
His reaction times were substantially faster in the World Track and
Field Championships compared to the United States Track and Field Championships, 
as were his teammates. 
This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction 
times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts.

@@

  \end{center}
\end{table}
The dataset of the four United States athletes 
is too small to perform any meaningful statistical analysis, but
the data was expanded to include reaction times for athletes who competed at 
national competitions (many of which were held from May-July of 2022). Once the 
data was collected, each athlete was designated as a cluster and the stage of
competition (national or international) was specified for each competition 
within the cluster (athlete).

\eds{Can you show some descriptive statistics for this data?  Total number of 
clusters? Average number of international vs national competitions across 
athletes/clusters?}
\eds{The following paragraph still needs to be smoothed (the paragraph that was 
here and the paragraph that was in the intro still need to be integrated into 
one (or two) cohesive paragraph(s) that accurately describe the two types of 
data that will be used for two rank-based analyses.}

\subsection{2019 World Track and Field Championships}
The 2019 World Track and Field Championship data is included in the larger dataset
described in section~\ref{sec:databeyond}, but we can also use it similarly to
how we used national level in section~\ref{sec:datanational}.  We can compare
the times of athletes who participated in the 2019 and 2022 World Track and Field
Championships to see if there were lower times in 2022.  While it is theoretically
possible that athletes got faster between 2019 and 2022, it is important to
remember that these are some of the fastest sprinters on the planet, and thus
improving on elite times is not easy.  Nonetheless, this is a worthy inclusion
to the data because Seiko was the timing company responsible for both events,
the World Championships have the biggest stakes, and because

First we can compare the reaction times of 
athletes who
competed at the World Track and Field Championships in both 2019 and 2022, and 
then we can compare how athletes performed at the 2022 championships relative 
to their national/qualifiers meet. In order to reduce any chances
of athletes improving reaction time, we can also look at data from qualifying 
meets. Almost every athlete had to qualify by proving they are elite at the 
national level, usually by coming in at least fourth. Thus, data was compiled on 
athletes who competed at the 2022 World Track and Field Championships on what 
these athletes also ran at their country's national championship. If there was 
any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August. The equipment used to measure 
reaction time cannot be guaranteed to be the same from national to international
competition, but the time span between meets has been reduced considerably.

\begin{figure}[tbp]
  \centering
  \includegraphics{Scatterplot2019vs2022}
  \caption{The reaction times from athletes who competed at least once in both 
  the 2019 and 2022 World Track and Field Championships}
  \label{Scatterplot2019vs2022}
\end{figure}
\of{I included this here for now, but feel free to change the location.  I will
add references to it and explanation of it along with some descriptive statistics about
the size and mean/median,std dev, size of the data.}
\jy{Show dot plots which would give the distribution; without uncertainty
  measure, the mean does not tell much.}
